Reasoning About a Machine with Local Capabilities

### 2018-04-20 toplas/paper.tex
added toplas directory and tex files with suggestions for what to add

%% For double-blind review submission
\documentclass{llncs}
%% For single-blind review submission
%\documentclass[acmsmall,10pt,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[acmsmall10pt,]{acmart}\settopmatter{}

%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format should change 'acmsmall' to
%% 'sigplan'.

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

@@

\newcommand{\sectionname}{Section}
\renewcommand{\figurename}{Figure}
\newcommand{\forcenewline}{$\phantom{v}$\\}
\newcommand{\judgment}[2]{\paragraph{#1}\hspace{\stretch{1}}\fbox{$#2$}}

@@

@@

\bibliographystyle{splncs}

% Authors and affiliation
\author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
\institute{Aarhus University \\ \email{\{lau,birkedal\}@cs.au.dk}
\and imec-DistriNet, KU Leuven \\ \email{dominique.devriese@cs.kuleuven.be}}

@@

Compromising software security is often based on attacks that break programming
language properties relied upon by software authors, such as control-flow
correctness, local-state encapsulation, etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control-flow correctness
and local-state encapsulation are either limited to only certain common forms of
attacks (leading to an attack-defense arms race) and/or rely on techniques like
machine code rewriting~\cite{wahbe_efficient_1993,abadi_control-flow_2005}, machine code
verification~\cite{morrisett_system_1999}, virtual machines with a
native stack~\cite{lindholm_java_2014} or
randomization~\cite{forrest_building_1997}. The latter techniques essentially
emulate protection techniques on existing hardware, at the cost of performance,
system complexity and/or security.

\emph{Capability machines} are a type of processors that
remediate these limitations with a better security model at
the hardware level. They are based on old ideas~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999},
but have recently received renewed interest; in particular, the CHERI
project has proposed new ideas and ways of tackling practical
challenges like backwards compatibility and realistic OS
support~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
machines tag every word (in the register file and in memory) to
enforce a strict separation between numbers and capabilities (a kind
of pointers that carry authority). Memory capabilities carry
the authority to read and/or write to a range of memory
locations. There is also a form of \emph{object capabilities}, which represent the
authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g., the M-Machine's enter capabilities or
CHERI's sealed code/data pairs).

@@

A similar problem is the enforcement of control-flow correctness on capability
machines. An interesting approach is taken in CheriBSD~\cite{Watson2015Cheri}:
the standard contiguous C stack is split into a central, trusted stack, managed
by trusted call and return instructions, and disjoint, private, per-compartment
stacks. To prevent illegal use of stack references, the approach relies on
\emph{local capabilities}, a type of capabilities offered by CHERI to
\emph{temporarily} relinquish authority, namely for the duration of a function
invocation whereafter the capability can be revoked. However, details are scarce
(how does it work precisely? what features are supported?) and a lot remains to
be investigated (e.g., combining disjoint stacks with cross-domain function
pointers seems like it will scale poorly to large numbers of components?).
Finally, there is no argument that the approach is watertight and it is not even
clear what security property is targeted exactly.

In this paper, we make two main contributions: (1) an alternative calling
convention that uses local capabilities to enforce stack frame encapsulation and
well-bracketed control flow, and (2) perhaps more importantly, we adapt and
apply the well-studied techniques of step-indexed Kripke logical relations for
reasoning about code on a representative capability machine with local
capabilities in general and correctness and security of the calling convention
in particular. More specifically, we make the following contributions:
\begin{itemize}
\item We formalize a simple but representative capability machine featuring local
  capabilities and its operational semantics
  (\sectionname~\ref{sec:capab-mach-with}).
\item We define a novel calling convention enforcing control-flow correctness and
  encapsulation of stack frames (\sectionname~\ref{sec:stack-and-return-pointer}). It
  relies solely on local capabilities and does not require OS support (like a
  trusted stack or call/return instructions).
  %It also does not use separate per-component stacks but a single, fairly standard, contiguous stack.
  %domi: don't mention this here. I think some CSF reviewers misunderstood that we were claiming this as a strength of the approach
  It supports higher-order cross-component calls (e.g., cross-component function
  pointers) and can be efficient assuming only one additional piece of processor
  support: an efficient instruction for clearing a range of memory.
\item We present a novel step-indexed Kripke logical relation for reasoning
  about programs on the capability machine. It is an untyped logical relation,
  inspired by previous work on object capabilities \cite{Devriese:2016ObjCap}.
  We prove an analogue of the standard fundamental theorem of logical relations
  --- to the best of our knowledge, our theorem is % by far
  the most general and powerful formulation of the formal guarantees offered by
  a capability machine (a form of capability
  safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
  guarantees offered for local capabilities. It is very general and not tied to
  our calling convention or a specific way of using the system's capabilities.
  We are the first to apply these techniques for reasoning about capability
  machines and we believe they will prove useful for many other purposes than
  our calling convention.
\item We introduce two novel technical ideas in the unary, step-indexed Kripke
  logical relation used to formulate the above theorem: the use of a
  \emph{single} orthogonal closure (rather than the earlier used biorthogonal
  closure) and a variant of Dreyer et al.~\cite{Dreyer:jfp12}'s public and private future
  worlds~\cite{Dreyer:jfp12} to express the special nature of local
  capabilities. The logical relation and the fundamental theorem expressing
  capability safety are presented in \sectionname~\ref{sec:logical-relation}.
\item We demonstrate our results by applying them to challenging examples,
  specifically constructed to demonstrate local-state encapsulation and
  control-flow correctness guarantees in the presence of cross-component
  function pointers (\sectionname~\ref{sec:examples}). The examples demonstrate both
  the power of our formulation of capability safety and our calling convention.
\end{itemize}
%\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution. Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}

For reasons of space, some details and all proofs have been omitted; please
refer to the technical appendix~\cite{technical_appendix} for
those.\dominique{upload TA as supplementary material instead?}

@@

\section{A Capability Machine with Local Capabilities}
\label{sec:capab-mach-with}
In this paper, we work with a formal capability machine with all the
characteristics of real capability machines, as well as local capabilities much
like CHERI's. Otherwise, it is kept as simple as possible. It is inspired by
both the M-Machine~\cite{Carter:1994:HSF:195473.195579} and
CHERI~\cite{Watson2015Cheri}. To avoid uninteresting details, we assume an
infinite address space and unbounded integers.

@@

We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
We assume an infinite set of addresses $\Addrs$ and define machine words as
either integers or capabilities of the form
$((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability represents the
authority to execute permissions $\perm$ on the memory range
$[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag
$\gl$ indicating whether the capability is global or local. There is no notion
of pointers other than capabilities, so we will use the terms interchangeably.
The available permissions and their ordering are depicted in
\figurename~\ref{fig:perm-hier}: the permissions include null permission
($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute
($\exec$) and read/write/execute ($\rwx$) permissions. Additionally, there are
three special permissions: read/write-local ($\readwritel$), read/write-local/execute
($\rwlx$) and enter ($\entry$), which we will explain below.

@@

\begin{comment}
  \begin{figure}
    \centering
    \begin{mathpar}
      \Phi \rightarrow
      \begin{cases}
        \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
        \begin{array}[t]{l}
          \text{if $\memreg(\pcreg) = \stdcap$}\\
          \;\;\text{and $\start \leq \addr \leq \addrend$ and $\memheap(\addr) = n$}\\
          \;\;\text{and $\perm \in \{ \exec,\rwx, \rwlx \}$} 
        \end{array}\\
        \failed                & \text{otherwise}
      \end{cases}
      \and \updatePcPerm{w} =
      \begin{cases}
        ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
        w & \text{otherwise}
      \end{cases}\\
      \and \stdUpdatePc{\Phi} =
      \begin{cases}
        \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
        \begin{array}[t]{l}
          \text{if $\memreg(\pcreg) = \stdcap$}\\ 
          \;\;\text{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}
        \end{array}\\
        \failed & \text{otherwise}
      \end{cases}
      \and \inferrule{ } { \sem{\fail}(\Phi) = \failed } \and
      \inferrule{ } { \sem{\halt}(\Phi) = (\halted,\memheap) } \and
      \inferrule{ } { \sem{\move{\refreg{r_1}}{r_2}}(\Phi) =
        \stdUpdatePc{}(\updateReg{r_1}{\memreg(r_2)}) } \and
      \inferrule{ \memreg(r_2) = \stdcap \\
        \perm \in \{ \rwx, \rwlx, \exec, \readwrite,
        \readwritel, \readonly \}\\
        \start \leq \addr \leq \addrend } {
        \sem{\load{\refreg{r_1}}{\refheap{r_2}}}(\Phi) =
        \updateReg{r_1}{\memheap(\addr)}} \and
      \inferrule{ \memreg(r_2) = \stdcap[\permp]\\
        \permp' = \decodePermPair{\memreg(r_2)}\\
        \permp'\sqsubseteq \permp } {
        \sem{\restricttwo{\refreg{r_1}}{r_2}}(\Phi) = \\
        \stdUpdatePc{}(\updateReg{r_1}{(\permp',\start,\addrend,\addr)})
      } \and \inferrule{ \memreg(r_2) = ((\_,\_),\_,\_,\addr) } {
        \sem{\geta{\refreg{r_1}}{\refreg{r_2}}}(\Phi) =
        \stdUpdatePc{\updateReg{r_1}{\addr}}} \and \inferrule{
        \var{newPc} = \updatePcPerm{\memreg(\lv)} } {
        \sem{\jmp{\lv}}(\Phi) = \updateReg{\pcreg}{\var{newPc}} } \and
      \inferrule{ \memreg(r_1) = \stdcap\\
        \start \leq \addr \leq \addrend\\
        \perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}\\
        \var{w} = \memreg(r_2) \\
        \var{w} = ((\_,\local),\_,\_,\_)\Rightarrow\perm \in
        \{\rwlx,\readwritel \} } {
        \sem{\store{\refheap{r_1}}{\refreg{r_2}}}(\Phi) =
        \stdUpdatePc{\updateHeap{\addr}{\var{w}}} } \and
    \end{mathpar}
    \caption{An excerpt from the operational semantics.}
    \label{fig:op-sem}
  \end{figure}
\end{comment}
\figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics for a
few representative instructions. Essentially, a configuration $\Phi$ either
decodes and executes the instruction at $\memreg(\pcreg)$ if it is executable
and its address is in the valid range or otherwise fails. The table in the
figure shows for instructions $i$ the result of executing them in configuration
$\Phi$. $\fail$ and $\halt$ obviously fail and halt respectively. $\move{}{}$
simply modifies the register file as requested and updates the $\pcreg$ to the
next instruction using the meta-function $\stdUpdatePc{}$.

The $\load{}{}$ instruction loads the contents of the requested memory location
into a register, but only if the capability has appropriate authority (i.e.\
read permission and an appropriate range). $\restricttwo{}{}$ updates a
capability's permissions and global/local tag in the register file, but only if
the new permissions are weaker than the original. It also never turns local
capabilities into global ones. $\geta{}{}$ queries the current address of a
capability and stores it in a register.

The $\jmp{}$ instruction updates the program counter to a requested location,
but it is complicated by the presence of \emph{enter capabilities}, modeled
after the M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
cannot be used to read, write or execute and their address and range cannot be
modified. They can only be used to jump to, but when that happens, their
permission changes to $\exec$. They can be used to represent a kind of closures:
an opaque package containing a piece of code together with local encapsulated
state. Such a package can be built as an enter capability $c =
((\entry,\gl),\start,\addrend,\addr)$ where the range $[\start,\addr-1]$
contains local state (data or capabilities) and $[\addr,\addrend]$ contains
instructions. The package is opaque to an adversary holding $c$ but when $c$ is
jumped to, the instructions can start executing and have access to the local
data through the updated version of $c$ that is then in $\pcreg$. 
\lau{As we talked about what is described here, is not how we make closures in
the TR, so we should consider whether we should refrain from using that word
here. }

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities}, modeled
after the ones in the CHERI processor~\cite{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we take care to clear the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
where register contents can be stored, including local capabilities. As long as
all capabilities with write-local authority are themselves local and the stack
is cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

@@

In order to have a reasonably realistic system, we use a simple model of linking
where a program has access to a linking table that contains capabilities for
other programs.  We also assume malloc to be part of the trusted computing base
satisfying a certain specification. Malloc and linking tables are described
further in the next section, but we refer to the technical
appendix~\cite{technical_appendix} for full details.

\section{Stack and Return Pointer Management Using Local Capabilities}
\label{sec:stack-and-return-pointer} One of the contributions in this paper is a
demonstration that local capabilities on a capability machine support a calling
convention that enforces control-flow correctness in a way that is provably
watertight, potentially efficient, does not rely on a trusted central stack
manager and supports higher-order interfaces to an adversary, where an adversary
is just some unknown piece of code. In this section, we explain this
convention's high-level approach, the security measures to be taken in a number
of situations (motivating each separately with a summary table at the end).
After that, we define a number of reusable macro-instructions that can be used
to conveniently apply the proposed convention in subsequent examples.

>>>>

\documentclass[format=acmsmall, review=false, screen=true]{acmart}

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs

% Metadata Information
\acmJournal{TWEB}
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2010}
\acmMonth{3}
\copyrightyear{2009}
%\acmArticleSeq{9}
% Copyright
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{0000001.0000001}

% Paper history
%\received{February 2007}
%\received[revised]{March 2009}
%\received[accepted]{June 2009}

\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

@@

\renewcommand{\figurename}{Figure}
\newcommand{\forcenewline}{$\phantom{v}$\\}
\newcommand{\judgment}[2]{\paragraph{#1}\hspace{\stretch{1}}\fbox{$#2$}}

@@

\definecolor{OliveGreen}{RGB}{60,128,49}
\newenvironment{toplas}
    {\color{OliveGreen}

    }{}
\newcommand{\itoplas}[1]
    {{\color{OliveGreen} #1}}

@@

\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}

% Authors and affiliation
%% Author with single affiliation.
\author{Lau~Skorstengaard}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
  \department{Department of Computer Science}              %% \department is recommended
  \institution{Aarhus University}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{Aarhus}
  \state{State1}
  \postcode{8210}
  \country{Denmark}
}
\email{lau@cs.au.dk}          %% \email is recommended

\author{Dominique~Devriese}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
  \department{iMinds-DistriNet}              %% \department is recommended
  \institution{KU Leuven}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{-}
  \state{-}
  \postcode{-}
  \country{Belgium}
}
\email{dominique.devriese@cs.kuleuven.be}          %% \email is recommended

\author{Lars~Birkedal}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
  \department{Department of Computer Science}              %% \department is recommended
  \institution{Aarhus University}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{Aarhus}
  \state{State1}
  \postcode{8210}
  \country{Denmark}
}
\email{birkedal@cs.au.dk}          %% \email is recommended
% \author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
% \institute{Aarhus University \\ \email{\{lau,birkedal\}@cs.au.dk}
% \and imec-DistriNet, KU Leuven \\ \email{dominique.devriese@cs.kuleuven.be}}

@@

Compromising software security is often based on attacks that break programming
language properties relied upon by software authors, such as control-flow
correctness, local-state encapsulation, etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control-flow correctness
and local-state encapsulation are either limited to only certain common forms of
attacks (leading to an attack-defense arms race) and/or rely on techniques like
machine code rewriting~\citet{wahbe_efficient_1993,abadi_control-flow_2005}, machine code
verification~\citet{morrisett_system_1999}, virtual machines with a
native stack~\citep{lindholm_java_2014} or
randomization~\citep{forrest_building_1997}. The latter techniques essentially
emulate protection techniques on existing hardware, at the cost of performance,
system complexity and/or security.

\emph{Capability machines} are a type of processors that
remediate these limitations with a better security model at
the hardware level. They are based on old ideas~\citet{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999},
but have recently received renewed interest; in particular, the CHERI
project has proposed new ideas and ways of tackling practical
challenges like backwards compatibility and realistic OS
support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
machines tag every word (in the register file and in memory) to
enforce a strict separation between numbers and capabilities (a kind
of pointers that carry authority). Memory capabilities carry
the authority to read and/or write to a range of memory
locations. There is also a form of \emph{object capabilities}, which represent the
authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g., the M-Machine's enter capabilities or
CHERI's sealed code/data pairs).

@@

A similar problem is the enforcement of control-flow correctness on capability
machines. An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}:
the standard contiguous C stack is split into a central, trusted stack, managed
by trusted call and return instructions, and disjoint, private, per-compartment
stacks. To prevent illegal use of stack references, the approach relies on
\emph{local capabilities}, a type of capabilities offered by CHERI to
\emph{temporarily} relinquish authority, namely for the duration of a function
invocation whereafter the capability can be revoked. However, details are scarce
(how does it work precisely? what features are supported?) and a lot remains to
be investigated (e.g., combining disjoint stacks with cross-domain function
pointers seems like it will scale poorly to large numbers of components?).
Finally, there is no argument that the approach is watertight and it is not even
clear what security property is targeted exactly.

In this paper, we make two main contributions: (1) an alternative calling
convention that uses local capabilities to enforce stack frame encapsulation and
well-bracketed control flow, and (2) perhaps more importantly, we adapt and
apply the well-studied techniques of step-indexed Kripke logical relations for
reasoning about code on a representative capability machine with local
capabilities in general and correctness and security of the calling convention
in particular. More specifically, we make the following contributions:
\begin{itemize}
\item We formalize a simple but representative capability machine featuring local
  capabilities and its operational semantics
  (\sectionname~\ref{sec:capab-mach-with}).
\item We define a novel calling convention enforcing control-flow correctness and
  encapsulation of stack frames (\sectionname~\ref{sec:stack-and-return-pointer}). It
  relies solely on local capabilities and does not require OS support (like a
  trusted stack or call/return instructions).
  %It also does not use separate per-component stacks but a single, fairly standard, contiguous stack.
  %domi: don't mention this here. I think some CSF reviewers misunderstood that we were claiming this as a strength of the approach
  It supports higher-order cross-component calls (e.g., cross-component function
  pointers) and can be efficient assuming only one additional piece of processor
  support: an efficient instruction for clearing a range of memory.
\item We present a novel step-indexed Kripke logical relation for reasoning
  about programs on the capability machine. It is an untyped logical relation,
  inspired by previous work on object capabilities \citet{Devriese:2016ObjCap}.
  We prove an analogue of the standard fundamental theorem of logical relations
  --- to the best of our knowledge, our theorem is % by far
  the most general and powerful formulation of the formal guarantees offered by
  a capability machine (a form of capability
  safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
  guarantees offered for local capabilities. It is very general and not tied to
  our calling convention or a specific way of using the system's capabilities.
  We are the first to apply these techniques for reasoning about capability
  machines and we believe they will prove useful for many other purposes than
  our calling convention.
\item We introduce two novel technical ideas in the unary, step-indexed Kripke
  logical relation used to formulate the above theorem: the use of a
  \emph{single} orthogonal closure (rather than the earlier used biorthogonal
  closure) and a variant of \citet{Dreyer:jfp12}'s public and private future
  worlds~\citep{Dreyer:jfp12} to express the special nature of local
  capabilities. The logical relation and the fundamental theorem expressing
  capability safety are presented in \sectionname~\ref{sec:logical-relation}.
\item We demonstrate our results by applying them to challenging examples,
  specifically constructed to demonstrate local-state encapsulation and
  control-flow correctness guarantees in the presence of cross-component
  function pointers (\sectionname~\ref{sec:examples}). The examples demonstrate both
  the power of our formulation of capability safety and our calling convention.
\end{itemize}
%\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution. Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}

For reasons of space, some details and all proofs have been omitted; please
refer to the technical appendix~\citep{technical_appendix} for
those.\dominique{upload TA as supplementary material instead?}

@@

\section{A Capability Machine with Local Capabilities}
\label{sec:capab-mach-with}
In this paper, we work with a formal capability machine with all the
characteristics of real capability machines, as well as local capabilities much
like CHERI's. Otherwise, it is kept as simple as possible. It is inspired by
both the M-Machine~\citep{Carter:1994:HSF:195473.195579} and
CHERI~\citep{Watson2015Cheri}. To avoid uninteresting details, we assume an
infinite address space and unbounded integers.

@@

We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
We assume an infinite set of addresses $\Addrs$ and define machine words as
either integers or capabilities of the form
$((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability represents the
authority to execute permissions $\perm$ on the memory range
$[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag
$\gl$ indicating whether the capability is global or local. There is no notion
of pointers other than capabilities, so we will use the terms interchangeably.
The available permissions and their ordering are depicted in
\figurename~\ref{fig:perm-hier}: the permissions include null permission
($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute
($\exec$) and read/write/execute ($\rwx$) permissions. Additionally, there are
three special permissions: read/write-local ($\readwritel$), read/write-local/execute
($\rwlx$) and enter ($\entry$), which we will explain below.
\itoplas{Suggestion: add the locality hierarchy and a short description. Maybe introduce the $\sqsubseteq$ relation here.}

@@

\begin{comment}
  \begin{figure}
    \centering
    \begin{mathpar}
      \Phi \rightarrow
      \begin{cases}
        \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
        \begin{array}[t]{l}
          \text{if $\memreg(\pcreg) = \stdcap$}\\
          \;\;\text{and $\start \leq \addr \leq \addrend$ and $\memheap(\addr) = n$}\\
          \;\;\text{and $\perm \in \{ \exec,\rwx, \rwlx \}$} 
        \end{array}\\
        \failed                & \text{otherwise}
      \end{cases}
      \and \updatePcPerm{w} =
      \begin{cases}
        ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
        w & \text{otherwise}
      \end{cases}\\
      \and \stdUpdatePc{\Phi} =
      \begin{cases}
        \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
        \begin{array}[t]{l}
          \text{if $\memreg(\pcreg) = \stdcap$}\\ 
          \;\;\text{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}
        \end{array}\\
        \failed & \text{otherwise}
      \end{cases}
      \and \inferrule{ } { \sem{\fail}(\Phi) = \failed } \and
      \inferrule{ } { \sem{\halt}(\Phi) = (\halted,\memheap) } \and
      \inferrule{ } { \sem{\move{\refreg{r_1}}{r_2}}(\Phi) =
        \stdUpdatePc{}(\updateReg{r_1}{\memreg(r_2)}) } \and
      \inferrule{ \memreg(r_2) = \stdcap \\
        \perm \in \{ \rwx, \rwlx, \exec, \readwrite,
        \readwritel, \readonly \}\\
        \start \leq \addr \leq \addrend } {
        \sem{\load{\refreg{r_1}}{\refheap{r_2}}}(\Phi) =
        \updateReg{r_1}{\memheap(\addr)}} \and
      \inferrule{ \memreg(r_2) = \stdcap[\permp]\\
        \permp' = \decodePermPair{\memreg(r_2)}\\
        \permp'\sqsubseteq \permp } {
        \sem{\restricttwo{\refreg{r_1}}{r_2}}(\Phi) = \\
        \stdUpdatePc{}(\updateReg{r_1}{(\permp',\start,\addrend,\addr)})
      } \and \inferrule{ \memreg(r_2) = ((\_,\_),\_,\_,\addr) } {
        \sem{\geta{\refreg{r_1}}{\refreg{r_2}}}(\Phi) =
        \stdUpdatePc{\updateReg{r_1}{\addr}}} \and \inferrule{
        \var{newPc} = \updatePcPerm{\memreg(\lv)} } {
        \sem{\jmp{\lv}}(\Phi) = \updateReg{\pcreg}{\var{newPc}} } \and
      \inferrule{ \memreg(r_1) = \stdcap\\
        \start \leq \addr \leq \addrend\\
        \perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}\\
        \var{w} = \memreg(r_2) \\
        \var{w} = ((\_,\local),\_,\_,\_)\Rightarrow\perm \in
        \{\rwlx,\readwritel \} } {
        \sem{\store{\refheap{r_1}}{\refreg{r_2}}}(\Phi) =
        \stdUpdatePc{\updateHeap{\addr}{\var{w}}} } \and
    \end{mathpar}
    \caption{An excerpt from the operational semantics.}
    \label{fig:op-sem}
  \end{figure}
\end{comment}
\figurename~\ref{fig:op-sem} \itoplas{Suggestion: extend this figure with more (all?) of the operational semantics. In general, should we add all the missing definitions and make this document closer to being self-contained.} 
shows an excerpt of the operational semantics for a
few representative instructions. Essentially, a configuration $\Phi$ either
decodes and executes the instruction at $\memreg(\pcreg)$ if it is executable
and its address is in the valid range or otherwise fails. The table in the
figure shows for instructions $i$ the result of executing them in configuration
$\Phi$. $\fail$ and $\halt$ obviously fail and halt respectively. $\move{}{}$
simply modifies the register file as requested and updates the $\pcreg$ to the
next instruction using the meta-function $\stdUpdatePc{}$.

The $\load{}{}$ instruction loads the contents of the requested memory location
into a register, but only if the capability has appropriate authority (i.e.\
read permission and an appropriate range). $\restricttwo{}{}$ updates a
capability's permissions and global/local tag in the register file, but only if
the new permissions are weaker than the original
%\itoplas{, which is what the relation $\sqsubseteq$ encapsulates.}
.
\itoplas{Reviewer E, popl, would like $\sqsubseteq$ and $\decodePermPair{}$ explained. If we explain all of these details, this section becaomes a bit long and cluncky. (see tex for suggestion)}
%\itoplas{As the permissions and global/local tag are not words on the machine, we decode the word in register $r_2$ as a pair of permission and locality which is done with the function $\decodePermPair{}$.}
% Reviewer E, popl: Figure 3: how is decodePermPair defined, and what is the relation between (perm', g') and (perm, g)?
It also never turns local
capabilities into global ones. $\geta{}{}$ queries the current address of a
capability and stores it in a register.

The $\jmp{}$ instruction updates the program counter to a requested location,
but it is complicated by the presence of \emph{enter capabilities}, modeled
after the M-Machine's~\citep{Carter:1994:HSF:195473.195579}. Enter capabilities
cannot be used to read, write or execute and their address and range cannot be
modified. They can only be used to jump to, but when that happens, their
permission changes to $\exec$. They can be used to represent a kind of closures:
an opaque package containing a piece of code together with local encapsulated
state. Such a package can be built as an enter capability $c =
((\entry,\gl),\start,\addrend,\addr)$ where the range $[\start,\addr-1]$
contains local state (data or capabilities) and $[\addr,\addrend]$ contains
instructions. The package is opaque to an adversary holding $c$ but when $c$ is
jumped to, the instructions can start executing and have access to the local
data through the updated version of $c$ that is then in $\pcreg$. 
\lau{As we talked about what is described here, is not how we make closures in
the TR, so we should consider whether we should refrain from using that word
here. }

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities}, modeled
after the ones in the CHERI processor~\citep{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we take care to clear the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
where register contents can be stored, including local capabilities. As long as
all capabilities with write-local authority are themselves local and the stack
is cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

@@

In order to have a reasonably realistic system, we use a simple model of linking
where a program has access to a linking table that contains capabilities for
other programs.  We also assume malloc to be part of the trusted computing base
satisfying a certain specification. Malloc and linking tables are described
further in the next section, but we refer to the technical
appendix~\citep{technical_appendix} for full details.
\itoplas{Suggestion: add the malloc specification (maybe later as it uses the LR) and possibly a description of the linking tables. (Reviewer A, esop suggest explaining the linking model)}
\itoplas{Suggestion: We also have the flag table ussed for assertions, maybe we should add a description of it here.}

\itoplas{Reviewer C, popl, would have liked a discussion of the relationship between this capability machines and old capability machines from the litterature (Cambridge CAP \& IBM System 39, see tex comment).}
 % * I'm old, so I'd appreciate a brief discussion LCM's relationship
 %  with the capability hardware I (and perhaps other older readers, or
 %  those outside this precise subfield) are likely to have heard of ---
 %  the Cambridge CAP & IBM System 39.  I think the capabilities of the
 %  machine are not that far from CAP, but this machine has the local
 %  capabilities support. Is that right?

\section{Stack and Return Pointer Management Using Local Capabilities}
\label{sec:stack-and-return-pointer}
\itoplas{Suggestion: say something about arguments in this paragraph instead of defering it to the discussion only?}
\itoplas{Suggestion: Make it more clear that this CC can be used to protect multiple components (reviewer A, ESOP), see tex comment.}
\itoplas{Suggestion: Add some of our presentation illustration figures to illustrate some of the issues? May not be feasible as they rely on ``animation''.}
% ## Clarify attacker model earlier
% The proposed calling convention seems able to protect multiple
% components from each other, yet most of the paper discusses this
% protection only in terms of one single statically known
% "adversary". This seems to be selling short the work, and raising
% questions of whether this calling convention is much weaker
% security-wise than the CHERI one, which is clearly about protecting
% multiple components. It would be good to explain this better upfront,
% not just in the discussions section at the end.
\itoplas{Suggestion: Include a description of (possibly definition for) well-bracketedness? (requested by reviewer C, popl). Reviewer D, popl, asks for the same but also a definition of local state encapsulation. In the POPL notes, I wrote that I talked with Lars about it and this is something that is not easily defined. Perhaps we should have a discussion of this.}
One of the contributions in this paper is a
demonstration that local capabilities on a capability machine support a calling
convention that enforces control-flow correctness in a way that is provably
watertight, potentially efficient, does not rely on a trusted central stack
manager and supports higher-order interfaces to an adversary, where an adversary
is just some unknown piece of code. In this section, we explain this
convention's high-level approach, the security measures to be taken in a number
of situations (motivating each separately with a summary table at the end).
After that, we define a number of reusable macro-instructions that can be used
to conveniently apply the proposed convention in subsequent examples.
